---
title: "2/26/19 notes"
author: "Dom-Romanello"
date: "2/26/2019"
output: html_document
---

for homework search help: dpois

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

discrete (e.g., bernoulli, dice) vs continuous (e.g., poisson, monkey scratch) 

```{r}

a <- 4
b <- 8
x <- seq(from = a - (b - a), to = b + (b - a), by = 0.01)
fx <- dunif(x, min = a, max = b)  # dunif() evaluates the density at each x
plot(x, fx, ylim = c(0, max(fx) + 0.1), type = "l", xlab = "x", ylab = "f(x)", 
    main = "Probability Density Function")

```

bernoulli distribution special case of binomial distribution

```{r}
n <- 6  # number of trials
k <- 6  # exact number of successes
p <- 1/6
prob <- (factorial(n)/(factorial(k) * factorial(n - k))) * (p^k) * (1 - p)^(n - 
    k)
prob
```

```{r}
k <- 3  # exact number of successes
prob <- (factorial(n)/(factorial(k) * factorial(n - k))) * (p^k) * (1 - p)^(n - 
    k)
prob
```

```{r}
dbinom(x = k, size = n, prob = p)
```
```{r}
probset <- dbinom(x = 0:6, size = 6, prob = 1/6)  # x is number of successes, size is number of trials
barplot(probset, names.arg = 0:6, space = 0, xlab = "outcome", ylab = "Pr(X = outcome)", 
    main = "Probability Mass Function")
```

```{r}
n <- 10  # number of trials
k <- 3  # exact number of successes
p <- 1/6
prob <- (factorial(n)/(factorial(k) * factorial(n - k))) * (p^k) * (1 - p)^(n - 
    k)
prob
```

```{r}
x <- 0:10
l = 3.5
probset <- dpois(x = x, lambda = l)
barplot(probset, names.arg = x, space = 0, xlab = "x", ylab = "Pr(X = x)", main = paste0("Probability Mass Function\nlambda = ", 
    l))
```


## hw good to go on question 1! just graph it. 

```{r}
a <- 4
b <- 8
x <- seq(from = a - (b - a), to = b + (b - a), by = 0.01)
fx <- dunif(x, min = a, max = b)  # dunif() evaluates the density at each x
plot(x, fx, ylim = c(0, max(fx) + 0.1), type = "l", xlab = "x", ylab = "f(x)", 
    main = "Probability Density Function")
```


```{r}
a <- 6
b <- 8
x <- seq(from = a - (b - a), to = b + (b - a), by = 0.01)
fx <- dunif(x, min = a, max = b)  # dunif() evaluates the density at each x
plot(x, fx, ylim = c(0, max(fx) + 0.1), type = "l", xlab = "x", ylab = "f(x)", 
    main = "Probability Density Function")
```

```{r}

V <- runif(1000,6,8)

mean(V)

var(V)

```

```{r}

library(manipulate)

manipulate(ggplot(data = data.frame(x = c(mu - 4 * sigma, mu + 4 * sigma)), 
    aes(x)) + stat_function(fun = dnorm, args = list(mean = mu, sd = sigma), 
    n = 1000) + xlab("x") + ylab("f(x)") + labs(title = "Exploring the Normal Distribution", 
    subtitle = paste0("Cumulative Probability Under Curve = ", round(pnorm(mu + 
        nsigma * sigma, mu, sigma) - pnorm(mu - nsigma * sigma, mu, sigma), 
        3))) + stat_function(fun = dnorm, xlim = c(mu - nsigma * sigma, mu + 
    nsigma * sigma), args = list(mean = mu, sd = sigma), n = 1000, geom = "area", 
    fill = "red", alpha = 0.5, color = "red"), mu = slider(-100, 100, initial = 0, 
    step = 10), sigma = slider(0, 30, initial = 5, step = 1), nsigma = slider(0, 
    4, initial = 0, step = 0.25))

manipulate(plot(seq(from = (mu - 4 * sigma), to = (mu + 4 * sigma), length.out = 1000), 
    pnorm(seq(from = (mu - 4 * sigma), to = (mu + 4 * sigma), length.out = 1000), 
        mean = mu, sd = sigma), type = "l", xlim = c(mu - 4 * sigma, mu + 4 * 
        sigma), xlab = "x", ylab = "f(x)", main = "Cumulative Probability"), 
    mu = slider(-100, 100, initial = 0, step = 10), sigma = slider(0, 30, initial = 5, 
        step = 1)) 

p <- pnorm(8, mean = 6, sd = 2) - pnorm(7, mean = 6, sd = 2)
p
```

```{r}


v <- rnorm(1000,3.5,4)

mean(v)

var(v)

sd(v)

hist(v)

hist(v, breaks = seq(from = -15, to = 20, by = 0.5), probability = TRUE)

```


```{r}

qqnorm(v, main = "Normal QQ plot random normal variables")
qqline(v, col = "gray")

p <- ppoints(length(v))
head(p)

tail(p)

theoretical_q <- qnorm(p)

observed_q <- quantile(v, ppoints(v))

plot(theoretical_q, observed_q, main = "QQ plot - Random Normal variable", xlab = "Theoretical Quantiles", 
    ylab = "Sample Quantiles")
```

2/28/19 

Probability mass- discrete

Probability density- within a range of values

standard normal distribution, mean 0, sd = 1

z scores use standard normal distributions

use set.seed to get the same sample every time

```{r}

set.seed(1)

x <- rnorm (1000000, 25, 5)

hist (x)

mean(x)

sd(x)

```

```{r}

x

y <- 5000

n <- 10

s <- NULL

k <- 5000  # number of samples
n <- 10  # size of each sample
s <- NULL  # dummy variable to hold each sample
for (i in 1:k) {
    s[[i]] <- sample(x, size = n, replace = FALSE)
}
head(s)

for (i in 1:k) m<- {
mean (s[[i]])}

m

for (i in 1:k) n<- {
sd (s[[i]])}

n

3.339711/sqrt(10)

```

